```{r loading libraries and functions, include = FALSE}
library(ggplot2)
library(psych); library(pastecs)
library(boot); 
library(QuantPsyc)
library(psych); library(pastecs)
library(lme4)
library(nlme)
library(effects)
library(nortest)
library(mediation)
library(car)
library(dplyr)
library(haven)
library(HardyWeinberg)
library(performance)

###diagnostic tool for linear models from roger mundry
source('.plot_diagnostics.R')
source('.diagnostics_dfbeta_cooks.R')
```
```{r load input data, echo=FALSE}
#input data.
data <- read_sav("/some/directoryPoster_2017/tables/2048_subjects_processed_during_10kin1day_stored_in_data_pt_life.sav")
data=data[,c("PseudoID","SIC")]

reward_vals<-read.table("/some/directoryPaper_2019/raw_data/rewardmetrics_2044.dat",
                        header=T)
glob_vals<-read.table("/some/directoryPaper_2019/raw_data/results_global.dat", header=T)

node_vals<-read.table("/some/directoryPaper_2019/raw_data/node_specific_results_strength.dat", header=T)

all_infos<-read.csv("/some/directory/Preprocessed/derivatives/PV168_A1_Pilot_subject_list_inclusion_exclusion29.1.19.csv")

genetics_taq1<-read.csv("/some/directory/Raw/Genetics_raw/2017_6_SNPs/SNP_6_coded_risk_n1882.csv")
genetics_fto<-read_sav("/some/directory/Raw/Genetics_raw/2015_33_SNPs/SNP_coded_risk_n5770_corrected_syntax_2019.sav")

#merge files
reward_vals_merged<-merge(reward_vals,data, by.x="PseudoID")
all_conn_vals<-merge(reward_vals_merged, glob_vals, by.x="PseudoID")
all_conn_vals<-merge(all_conn_vals,node_vals, by.x="PseudoID")
length(all_conn_vals$PseudoID)
#merge demographic
all_conn_vals<-merge(all_conn_vals, all_infos, by.x="SIC")
```

```{r select only participants aged below 60 years, echo=FALSE}
nrow(all_conn_vals[all_conn_vals$Age_all>=60,])
all_conn_vals_age=all_conn_vals[all_conn_vals$Age_all<60,]
nrow(all_conn_vals_age)
```

```{r add genetic data, echo=FALSE}
#add genetic data: N=399 have genetic data
select_genetics=merge(all_conn_vals_age, genetics_taq1, by.x="SIC")
select_genetics=merge(select_genetics, genetics_fto, by.x="SIC", by.y="SIC")
nrow(select_genetics)

#recode 
select_genetics$Taq1A_risk=1
select_genetics[select_genetics$rs1800497_risk==0,"Taq1A_risk"]=0
select_genetics$FTO_risk=as.factor(select_genetics$FTO_risk)
select_genetics$Taq1A_risk=as.factor(select_genetics$Taq1A_risk)
```

```{r add BMI data missing in two participants, echo=FALSE}
#BMI
anthro=readxl::read_xlsx("/some/directory/Raw/Questionnaires_raw/2018_01_PV168_Update/data/PV0168_D00074_NODUP.xlsx")
select_genetics_bmi=merge(select_genetics,anthro,by.x="SIC",by.y="BMI_S010061_SIC",all.x=T)

#exclude participants without bmi information (N=2)
select_genetics_bmi=select_genetics_bmi[!is.na(select_genetics_bmi$BMI_BMI),]
nrow(select_genetics_bmi)
```

```{r apply inclusion/exclusion previously n:17 were excluded, echo=FALSE}
###
#stroke: 1
table(select_genetics_bmi[,c("stroke_y_n")])
#lesions + befund: 7 + 4
table(select_genetics_bmi[,c("mri_lesion_num")])
table(select_genetics_bmi[,c("MRT_BefundBewertung_Bef")])
#cancer treatment: 6
table(select_genetics_bmi[,c("cancer_treatment_y_n")])
#epilepsy:5
table(select_genetics_bmi[,c("epilepsy_y_n")])
#parkinson: none
table(select_genetics_bmi[,c("parkinson_y_n")])
#ms:2
table(select_genetics_bmi[,c("ms_y_n")])
#central medication: strong: 27, indirect: 5
table(select_genetics_bmi[,c("Medi_central")])

#inclusion exclusion -> 46 exclusion
include=(select_genetics_bmi$stroke_y_n!=1&
          (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
         select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein'&
         select_genetics_bmi$cancer_treatment_y_n!=1&
         select_genetics_bmi$mri_tumors_num!="behandlungsbedürftig"&
         select_genetics_bmi$epilepsy_y_n!=1&
         select_genetics_bmi$ms_y_n!=1&
         select_genetics_bmi$parkinson_y_n!=1&
         select_genetics_bmi$Medi_central!="strong/direct"&
         select_genetics_bmi$Medi_central!="weak/indirect")
table(include)

#Exclude home many for what
table(select_genetics_bmi$stroke_y_n!=1) #1 stroke
table(select_genetics_bmi$stroke_y_n!=1& # 7 mri lesions
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital"))
table(select_genetics_bmi$stroke_y_n!=1& #4 befund
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
          select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein')
table(select_genetics_bmi$stroke_y_n!=1& #5 cancer treatment
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
         select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein'&
         select_genetics_bmi$cancer_treatment_y_n!=1)
table(select_genetics_bmi$stroke_y_n!=1& #5 epilepsy
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
         select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein'&
         select_genetics_bmi$cancer_treatment_y_n!=1&
        select_genetics_bmi$epilepsy_y_n!=1)
table(select_genetics_bmi$stroke_y_n!=1& #1 ms
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
         select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein'&
         select_genetics_bmi$cancer_treatment_y_n!=1&
        select_genetics_bmi$epilepsy_y_n!=1&
        select_genetics_bmi$ms_y_n!=1)
table(select_genetics_bmi$stroke_y_n!=1& #25 cm
         (select_genetics_bmi$mri_lesion_num=="missing"|
          select_genetics_bmi$mri_lesion_num=="keine"|
          select_genetics_bmi$mri_lesion_num=="kongenital")&
         select_genetics_bmi$MRT_BefundBewertung_Bef!='verwendungsfähig nein'&
         select_genetics_bmi$cancer_treatment_y_n!=1&
        select_genetics_bmi$epilepsy_y_n!=1&
        select_genetics_bmi$ms_y_n!=1&
         select_genetics_bmi$Medi_central!="strong/direct"&
         select_genetics_bmi$Medi_central!="weak/indirect")


#N=349 von N=397 (48 excluded due to all of the above, out of which 25 due to centrally active medication)
tmp_sample=select_genetics_bmi[include,]

```



### Confounding variables:
```{r add confounders, echo=FALSE}
#add mean FD from DWI scan.
headmotion=read.csv("/some/directory/Preprocessed/derivatives/head_motion/LIFE3T_DWI.csv")
tmp_sample=merge(tmp_sample,headmotion,by.x="SIC",by.y="subject_id",all.x=T)

#smoking status:
sm=readxl::read_xlsx("/some/directory/Raw/Questionnaires_raw/2018_01_PV168_Update/data/PV0168_D00141_NODUP.xlsx")
tmp_sample=merge(tmp_sample,sm,by.x="SIC",by.y="TOB2_R00001_SIC",all.x=T)

#CESD
cesd=readxl::read_xlsx("/some/directory/Raw/Questionnaires_raw/2018_01_PV168_Update/data/PV0168_D00041_NODUP.xlsx")
tmp_sample=merge(tmp_sample,cesd,by.x="SIC",by.y="CES_D_SCORE_SIC",all.x=T)

#soziodemo
socio=readxl::read_xlsx("/some/directory/Raw/Questionnaires_raw/2018_01_PV168_Update/data/PV0168_T00001_NODUP.xlsx")
tmp_sample=merge(tmp_sample,socio,by.x="SIC",by.y="SOZIO_SIC",all.x=T)

#recode education information:
tmp_sample[!is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==95,"SOZIO_F0041"]=6

tmp_sample$educ_4levels=0
min(tmp_sample$SOZIO_F0041, na.rm=TRUE)
#tmp_sample[!is.na(tmp_sample$SOZIO_F0041)&
#                            tmp_sample$SOZIO_F0041==1|
#                            !is.na(tmp_sample$SOZIO_F0041)&
#                            tmp_sample$SOZIO_F0041==2,]$educ_4levels=1
tmp_sample[!is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==3|
                            tmp_sample$SOZIO_F0041==5&
                            tmp_sample$Age_all>=61,"educ_4levels"]=2
tmp_sample[!is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==4|
                            !is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==5&
                            tmp_sample$Age_all<61,"educ_4levels"]=3
tmp_sample[!is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==6|
                            !is.na(tmp_sample$SOZIO_F0041)&
                            tmp_sample$SOZIO_F0041==7,"educ_4levels"]=4
tmp_sample[tmp_sample$educ_4levels==0,"educ_4levels"]=NA
```
```{r select only relevant variables, echo=FALSE}
#select only relevant variables (to reduce the messiness)
colnames(tmp_sample)[c(1:40,43,44,73,75,77,80,86,117,118,138,141,237,448,453,490,493,497,498,500,505,511,514,608)]


final_sample=tmp_sample[,c(1:40,43,44,73,75,77,80,86,117,118,138,141,237,448,453,490,493,497,498,500,505,511,514,608)]
colnames(final_sample)

final_sample$BMI_z=scale(final_sample$BMI_BMI)
final_sample$age_z=scale(final_sample$Age_all)
final_sample$log_CESD=log(final_sample$CES_D_SCORE_SUM_CES_D+1)
final_sample$log_meanFD=log(final_sample$meanFD)
#levels(final_sample$Medi_central)=c("None","None","Strong","Weak")
```

```{r define outliers, echo=FALSE}
#according to 10-k-in-1-day workshop: mean FA/mean NOS < or > 
#Outliers were defined as values higher than (3rd quartile + 2*interquartile range (IQR))
#or lower than (1st quartile - 2*interquartile range (IQR)).
#exclude two participants

threshold_NOS=c(quantile(final_sample$mean_NOS)[2]-2*IQR(final_sample$mean_NOS),
             quantile(final_sample$mean_NOS)[4]+2*IQR(final_sample$mean_NOS))
threshold_FA=c(quantile(final_sample$mean_FA)[2]-2*IQR(final_sample$mean_FA),
                quantile(final_sample$mean_FA)[4]+2*IQR(final_sample$mean_FA))

final_sample[final_sample$mean_NOS<threshold_NOS[1],"SIC"]
final_sample[final_sample$mean_NOS>threshold_NOS[2],"SIC"]
final_sample[final_sample$mean_FA<threshold_FA[1],"SIC"]
final_sample[final_sample$mean_FA>threshold_FA[2],"SIC"]

final_sample=final_sample[final_sample$mean_FA<threshold_FA[2]&
                                                    final_sample$mean_FA>threshold_FA[1]&
                                                    final_sample$mean_NOS<threshold_NOS[2]&
                                                    final_sample$mean_NOS>threshold_NOS[1],]
nrow(final_sample)

#normalize values
final_sample$rel_FA_CC=final_sample$REW_FA_CC/final_sample$GLOB_FA_CC
final_sample$rel_FA_strength=final_sample$REW_FA_strength/final_sample$GLOB_FA_strength
final_sample$rel_NOS_CC=final_sample$REW_NOS_CC/final_sample$GLOB_NOS_CC
final_sample$rel_NOS_strength=final_sample$REW_NOS_strength/final_sample$GLOB_NOS_strength

#nodewise:
final_sample$rel_strength_NOS_l_latOFC=final_sample$strength_NOS_l_latOFC/final_sample$GLOB_NOS_strength      
final_sample$rel_strength_NOS_r_latOFC=final_sample$strength_NOS_r_latOFC/final_sample$GLOB_NOS_strength       
final_sample$rel_strength_NOS_l_mOFC=final_sample$strength_NOS_l_mOFC/final_sample$GLOB_NOS_strength
final_sample$rel_strength_NOS_r_mOFC=final_sample$strength_NOS_r_mOFC/final_sample$GLOB_NOS_strength         
final_sample$rel_strength_NOS_l_put=final_sample$strength_NOS_l_put/final_sample$GLOB_NOS_strength
final_sample$rel_strength_NOS_r_put=final_sample$strength_NOS_r_put/final_sample$GLOB_NOS_strength          
final_sample$rel_strength_NOS_l_caudate=final_sample$strength_NOS_l_caudate/final_sample$GLOB_NOS_strength
final_sample$rel_strength_NOS_r_caudate=final_sample$strength_NOS_r_caudate/final_sample$GLOB_NOS_strength      
final_sample$rel_strength_NOS_l_acc=final_sample$strength_NOS_l_acc/final_sample$GLOB_NOS_strength
final_sample$rel_strength_NOS_r_acc=final_sample$strength_NOS_r_acc/final_sample$GLOB_NOS_strength          
final_sample$rel_strength_FA_l_latOFC=final_sample$strength_FA_l_latOFC/final_sample$GLOB_FA_strength
final_sample$rel_strength_FA_r_latOFC=final_sample$strength_FA_r_latOFC/final_sample$GLOB_FA_strength       
final_sample$rel_strength_FA_l_mOFC=final_sample$strength_FA_l_mOFC/final_sample$GLOB_FA_strength
final_sample$rel_strength_FA_r_mOFC=final_sample$strength_FA_r_mOFC/final_sample$GLOB_FA_strength          
final_sample$rel_strength_FA_l_put=final_sample$strength_FA_l_put/final_sample$GLOB_FA_strength
final_sample$rel_strength_FA_r_put=final_sample$strength_FA_r_put/final_sample$GLOB_FA_strength           
final_sample$rel_strength_FA_l_caudate=final_sample$strength_FA_l_caudate/final_sample$GLOB_FA_strength
final_sample$rel_strength_FA_r_caudate=final_sample$strength_FA_r_caudate/final_sample$GLOB_FA_strength      
final_sample$rel_strength_FA_l_acc=final_sample$strength_FA_l_acc/final_sample$GLOB_FA_strength
final_sample$rel_strength_FA_r_acc=final_sample$strength_FA_r_acc/final_sample$GLOB_FA_strength
```

# DESCRIPTIVES
```{r create descriptives of sample, echo=FALSE}
colnames(final_sample)
final_sample$hypertension_composite_score[final_sample$hypertension_composite_score==999999] <- NA 

describe(final_sample[,c(41:63)])
table(final_sample$educ_4levels)
table(final_sample$FA_color_artifact_y_n)

write.csv(final_sample, "/some/directoryPaper_2019/scripts/final_sample.csv")
```

```{r Hardy-Weinberg, echo=FALSE}
table(tmp_sample[,"FTO_risk"])
MAF=(table(tmp_sample[,"FTO_risk"])[2]+2*table(tmp_sample[,"FTO_risk"])[3])/(nrow(tmp_sample[,])*2)
HW.test <- HWChisq(table(tmp_sample[,"FTO_risk"]),verbose=TRUE)

table(tmp_sample[,"rs1800497_risk"])
MAF=(table(tmp_sample[,"rs1800497_risk"])[2]+2*table(tmp_sample[,"rs1800497_risk"])[3])/(nrow(tmp_sample[,])*2)
HW.test <- HWChisq(table(tmp_sample[,"rs1800497_risk"]),verbose=TRUE)

```

# LINEAR MODELS
## FA CC ~ BMI (Model 1, Model 2 not done)
```{r FA CC model, echo=FALSE}
##FA_CC
R1 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z , data=final_sample)
R0 <- lm(rel_FA_CC ~ age_z + factor(sex), data=final_sample)
anova(R1,R0) #p=0.02419 -> no adjusted model
summary(R1)
#lm.beta(R1)
confint(R1)
cbind(coefficients(R1),confint(R1))

##graphically show confidence intervalls
round(confint(R1),4)
##first check assumptions
#diagnostics-> normally distributed residuals with homogenous variance
hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))

plot(x=fitted(R1),y=residuals(R1))
plot(residuals(R1))
abline(a=0,b=0)

qqnorm(residuals(R1))
qqline(residuals(R1))

##model stability-> looks ok
max(abs(dffits(R1)))
head(dfbeta(R1))
xx=cbind(coefficients(R1),coefficients(R1)+t(apply(dfbeta(R1),MARGIN=2, FUN=range)))
colnames(xx)=c("orig","min","max")
round(xx,5)
max(as.vector(influence(R1)$hat))

##MULTIcollinearity
qqPlot(R1)
vif(R1) # largest vif > 10 is problematic
1/vif(R1) # Tolerance < 0.1 serious problem; <0.2 potential problem
mean(vif(R1)) # Average SUBSTANTIALLY > 1, the regression might be biased
###
```

## FA CS ~ BMI (Model 1 +2)
```{r CS FA, echo=FALSE}
R1 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z , data=final_sample)
R0 <- lm(rel_FA_strength ~ age_z + factor(sex), data=final_sample)
anova(R1,R0) # 0.006171
summary(R1)
  confint(R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

R2.1=lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + as.factor(educ_4levels) +
          log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
          log_meanFD, data=final_sample)
R2.0=lm(rel_FA_strength ~ age_z + factor(sex) +  as.factor(educ_4levels) +
          log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
          log_meanFD, 
        data=final_sample)
anova(R2.1,R2.0)#0.01985
summary(R2.1)
confint(R2.1)

hist(residuals(R2.1),probability = T)
x=seq(from=min(residuals(R2.1)),
      to=max(residuals(R2.1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R2.1))))
qqPlot(R2.1)
```
## NOS CC ~ BMI (Model 1, Model 2 not calculated)
```{r NOS CC, echo=FALSE}
R1 <- lm(rel_NOS_CC ~ age_z + factor(sex)+ BMI_z , data=final_sample)
R0 <- lm(rel_NOS_CC ~ age_z + factor(sex), data=final_sample)
anova(R1,R0)#p=0.1969
summary(R1)
confint(R1)
```
## NOS CS ~ BMI (Model 1+2)
```{r NOS strength, echo=FALSE}
R1 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z, data=final_sample)
R0 <- lm(rel_NOS_strength ~ age_z + factor(sex) , data=final_sample)
anova(R0,R1)#p=0.002418
summary(R1)
confint(R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

R2.1=lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z + as.factor(educ_4levels) +
          log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
          log_meanFD, data=final_sample)
R2.0=lm(rel_NOS_strength ~ age_z + factor(sex) + as.factor(educ_4levels) +
          log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
          log_meanFD, 
        data=final_sample)
anova(R2.1,R2.0)#0.007409
summary(R2.1)
confint(R2.1)

hist(residuals(R2.1),probability = T)
x=seq(from=min(residuals(R2.1)),
      to=max(residuals(R2.1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R2.1))))
qqPlot(R2.1)

check_model(R2.1)
```
## Strength of individual nodes
```{r strength of individual node, echo=FALSE}
R1 <- lm(rel_strength_FA_l_acc ~ age_z + factor(sex) + BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_l_acc ~ age_z + factor(sex) + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.0026/0.026
summary(R1)
confint(R1)
R1 <- lm(rel_strength_FA_r_acc ~ age_z + factor(sex) + BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_r_acc ~ age_z + factor(sex) + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.00159/0.0043
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_l_caudate ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_l_caudate ~ age_z + factor(sex) + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.49/0.72
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_r_caudate ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_r_caudate ~ age_z + factor(sex) + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.15/0.12
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_l_put ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_l_put ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.0009/0.0019
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_r_put ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_r_put ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.007/0.017
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_l_mOFC ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_l_mOFC ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.081/0.082
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_r_mOFC ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_r_mOFC ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.27/0.44
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_l_latOFC ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD , data=final_sample)
R0 <- lm(rel_strength_FA_l_latOFC ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.62/0.63
summary(R1)
confint(R1)

R1 <- lm(rel_strength_FA_r_latOFC ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_FA_r_latOFC ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.0018/0.0055
summary(R1)
confint(R1)

###NOS
R1 <- lm(rel_strength_NOS_l_acc ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_l_acc ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p 0.05
summary(R1)
confint(R1)

R1 <- lm(rel_strength_NOS_r_acc ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_r_acc ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p 0.034
summary(R1)
confint(R1)
R1 <- lm(rel_strength_NOS_l_caudate ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_l_caudate ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p 0.16
summary(R1)
confint(R1)
qqplot(R1)
hist(final_sample$rel_strength_FA_l_caudate)

R1 <- lm(rel_strength_NOS_r_caudate ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_r_caudate ~ age_z + factor(sex)+  
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p=0.89
summary(R1)
confint(R1)
R1 <- lm(rel_strength_NOS_l_put ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_l_put ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p=0.0015
summary(R1)
confint(R1)

R1 <- lm(rel_strength_NOS_r_put ~ age_z + factor(sex)+ BMI_z + 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_r_put ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p=0.062
summary(R1)
confint(R1)
R1 <- lm(rel_strength_NOS_l_mOFC ~ age_z + factor(sex)+ BMI_z +
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_l_mOFC ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#adj p=0.024
summary(R1)
confint(R1)
R1 <- lm(rel_strength_NOS_r_mOFC ~ age_z + factor(sex)+ BMI_z +
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_r_mOFC ~ age_z + factor(sex)+ 
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.32
summary(R1)
confint(R1)

R1 <- lm(rel_strength_NOS_l_latOFC ~ age_z + factor(sex)+ BMI_z +
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD , data=final_sample)
R0 <- lm(rel_strength_NOS_l_latOFC ~ age_z + factor(sex)+
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.079
summary(R1)
confint(R1)

R1 <- lm(rel_strength_NOS_r_latOFC ~ age_z + factor(sex)+ BMI_z +
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
R0 <- lm(rel_strength_NOS_r_latOFC ~ age_z + factor(sex) +
           as.factor(educ_4levels) +
           log_CESD + as.factor(TOB2_SMOKING_STATUS_CURATED) + 
           log_meanFD, data=final_sample)
anova(R1,R0)#p=0.16
summary(R1)
confint(R1)
```

## GENETICS
### Genetic effects on BMI
```{r Genetic effects on BMI, echo=FALSE}
#trend level significance of interaction
R3 <- lm(BMI_BMI~ Age_all + factor(sex) + as.factor(FTO_risk)*as.factor(Taq1A_risk)  , data=final_sample)
R0 <- lm(BMI_BMI~ Age_all + factor(sex)+ as.factor(FTO_risk) + as.factor(Taq1A_risk), data=final_sample)
summary(R3)
anova(R3,R0)

#interpretation of main effects from this model:
R1 <- lm(BMI_BMI~ Age_all + factor(sex)+ as.factor(FTO_risk) + as.factor(Taq1A_risk), data=final_sample)
summary(R1)

#significance of FTO
R1 <- lm(BMI_BMI~ Age_all + factor(sex)+ as.factor(FTO_risk) + as.factor(Taq1A_risk), data=final_sample)
R0 <- lm(BMI_BMI~ Age_all + factor(sex) + as.factor(Taq1A_risk)  , data=final_sample)
anova(R1,R0)

#significance of Taq
R1 <- lm(BMI_BMI~ Age_all + factor(sex)+ as.factor(FTO_risk) + as.factor(Taq1A_risk), data=final_sample)
R0 <- lm(BMI_BMI~ Age_all + factor(sex) + as.factor(FTO_risk)  , data=final_sample)
anova(R1,R0)

```
```{r plot BMI and genotypes, echo=FALSE}
plot <- ggplot(final_sample, aes(x = as.factor(FTO_risk), y = BMI_BMI, colour = as.factor(Taq1A_risk)))+
        geom_boxplot(outlier.shape = NA) + 
        geom_jitter(position=position_jitterdodge(jitter.width=0.2, dodge.width=0.8), aes(colour = as.factor(Taq1A_risk))) +
        theme_classic()+
        ylab(bquote("BMI" ~ (kg/m^2))) + 
        xlab(bquote('# of FTO risk alleles')) +
        scale_color_discrete(name = "Taq1A risk allele", labels = c("no", "yes")) +
        theme(
        legend.position="top",
        axis.text = element_text(colour = "black"),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text( size=16),
        axis.title.y = element_text( size=16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16))#,
        #plot.margin=unit(c(1.5,0,0.4,1),"cm"))
plot  
  
ggsave("/some/directoryPaper_2019/manuscript/figures/Figure2_updated.jpeg", width = 185, height = 200, units = "mm", dpi = 300)  
ggsave("/some/directoryPaper_2019/manuscript/figures/Figure2_updated.eps", width = 185, height = 200, units = "mm", dpi = 300) 
ggsave("/some/directoryPaper_2019/manuscript/figures/Figure2_updated.tiff", width = 185, height = 200, units = "mm", dpi = 300)  
```

### FA CC
#### without BMI adjustment
```{r FA CC, echo =FALSE}
R1 <- lm(rel_FA_CC ~ age_z + factor(sex) + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
  R0 <- lm(rel_FA_CC ~ age_z + factor(sex) + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R1)
summary(R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

summary(R0)

R1 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + as.factor(Taq1A_risk) + as.factor(FTO_risk), data=final_sample)
anova(R0,R1)

summary(R0)
```
#### with BMI adjustment (used in paper, see above)
```{r FA CC, echo =FALSE}
R1 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_FA_CC ~ age_z + factor(sex), data=final_sample)
anova(R0,R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

R2 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + as.factor(Taq1A_risk)+as.factor(FTO_risk) , data=final_sample)
anova(R2,R0)
summary(R2)
confint(R2)
```
#### with all confounders  (not used in paper)
```{r FA CC, echo =FALSE}
R1 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD, data=final_sample)
anova(R0,R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

R2 <- lm(rel_FA_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk)+as.factor(FTO_risk) , data=final_sample)
anova(R2,R0)
```
### FA strength
#### without BMI adjustment
```{r FA strength without BMI, echo=FALSE}
R1 <- lm(rel_FA_strength ~ age_z + factor(sex)  + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_FA_strength ~ age_z + factor(sex) + Taq1A_risk +FTO_risk , data=final_sample)
anova(R1,R0)
summary(R1)

R2 <- lm(rel_FA_strength ~ age_z + factor(sex) + Taq1A_risk+FTO_risk, data=final_sample)
summary(R2)
confint(R2)

R1 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_FA_strength ~ age_z + factor(sex)+ BMI_z + Taq1A_risk +FTO_risk, data=final_sample)
anova(R1,R0)
summary(R0)
```
#### with BMI adjustment (not used for paper)
```{r FA strength with BMI, echo=FALSE}
R1 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_FA_strength ~ age_z + factor(sex)+ BMI_z, data=final_sample)
anova(R1,R0)
summary(R1)

R2 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + Taq1A_risk+FTO_risk, data=final_sample)
anova(R2,R0)
summary(R2)
```
#### with all confounders (not used for paper)
```{r FA CS, echo =FALSE}
R1 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD, data=final_sample)
summary(R1)
anova(R0,R1)

R2 <- lm(rel_FA_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
summary(R2)
anova(R0,R2)

```

### NOS CC 
#### without BMI adjustment
```{r NOS CC, echo=FALSE}
R1 <- lm(rel_NOS_CC ~ age_z + factor(sex) +  Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_NOS_CC ~ age_z + factor(sex) + Taq1A_risk+FTO_risk, data=final_sample)
anova(R1,R0)
summary(R0)
confint(R0)

R1 <- lm(rel_NOS_CC ~ age_z + factor(sex) +  BMI_z + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_NOS_CC ~ age_z + factor(sex) + BMI_z + Taq1A_risk +FTO_risk , data=final_sample)
anova(R1,R0)
summary(R0)
```
#### with BMI adjustment (see above)
```{r NOS CC, echo=FALSE}
R1 <- lm(rel_NOS_CC ~ age_z + factor(sex) +  BMI_z + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_NOS_CC ~ age_z + factor(sex) + BMI_z, data=final_sample)
anova(R1,R0)

R2 <- lm(rel_NOS_CC ~ age_z + factor(sex) +  BMI_z + Taq1A_risk+FTO_risk, data=final_sample)
anova(R2,R0)

```
#### with all adjustments (not used for paper)
```{r NOS CC, echo =FALSE}
R1 <- lm(rel_NOS_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_NOS_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD, data=final_sample)
summary(R1)
anova(R0,R1)

R2 <- lm(rel_NOS_CC ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R2)
summary(R2)
```

#NOS strength 
### without BMI adjustment
```{r NOS CS, echo=FALSE}
R1 <- lm(rel_NOS_strength ~ age_z + factor(sex) + Taq1A_risk*FTO_risk , data=final_sample)
R0 <- lm(rel_NOS_strength ~ age_z + factor(sex)+ Taq1A_risk+FTO_risk, data=final_sample)
anova(R1,R0)
summary(R1)
summary(R0)
confint(R0)

R1 <- lm(rel_NOS_strength ~ age_z + factor(sex) + Taq1A_risk*FTO_risk + BMI_z , data=final_sample)
R0 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z + Taq1A_risk+FTO_risk, data=final_sample)
anova(R1,R0)
summary(R0)
```
### with BMI adjustment
```{r NOS CS, echo=FALSE}
R1 <- lm(rel_NOS_strength ~ age_z + factor(sex) + Taq1A_risk*FTO_risk + BMI_z , data=final_sample)
R0 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z, data=final_sample)
anova(R1,R0)

R2 <- lm(rel_NOS_strength ~ age_z + factor(sex)+ BMI_z + Taq1A_risk+FTO_risk, data=final_sample)
anova(R2,R0)
summary(R2)

R3<- lm(rel_NOS_strength ~ age_z + factor(sex) +FTO_risk*BMI_z + Taq1A_risk, data=final_sample)
summary(R3)
anova(R3,R2)
```
### with all adjustments

```{r NOS CS, echo =FALSE}
R1 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
R0 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD, data=final_sample)
summary(R1)
anova(R0,R1)

R2 <- lm(rel_NOS_strength ~ age_z + factor(sex) + BMI_z + educ_4levels +
           log_CESD + TOB2_SMOKING_STATUS_CURATED + 
           log_meanFD + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R2)
summary(R2)
```
