---
title: "BMI, obesity-related genetic polymorphisms and reward network structural connectivity"
author: "Frauke Beyer. Markus Scholz, Rui Zhang, Kerstin Wirkner, Markus Loeffler, Michael Stumvoll, Arno Villringer, A. Veronica Witte "
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
---
```{r loading necessary libraries for R 3.6.1, include = FALSE}
library(ggplot2)
library(psych); 
library(pastecs)
library(boot); 
library(QuantPsyc)
library(lme4)
library(nlme)
library(effects)
library(nortest)
library(mediation)
library(car)
library(dplyr)
library(haven)
library(HardyWeinberg)
library(performance)
library(car)
library("gplots")
library(knitr)
library(kableExtra)
```

Here, you find the main analysis steps for the publication "	Higher BMI, but not obesity-related genetic polymorphisms, correlates with lower structural connectivity of the reward network in a population-based study" with a synthetic dataset based on the real data.
```{r load input data, echo=FALSE}
final_sample <- read.csv("synth_data.csv")
```

# Descriptives of the dataset
```{r create descriptives of sample, echo=FALSE}
tmp=describe(final_sample[,c("Age_all","BMI_BMI","rel_FA_CC","rel_FA_strength",
                         "rel_NOS_CC", "rel_NOS_strength")])

tmp %>%
   kable("html", escape = F, caption="Demographics of synthetic dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

table(final_sample[,"sex"])

cont_table=table(final_sample$FTO_risk,final_sample$Taq1A_risk)
dt <- as.table(cont_table)
balloonplot(t(dt), main="", xlab = "Taq1A risk allele carrier" , ylab="FTO_risk", show.margins = FALSE)
```
Check for Hardy-Weinberg equilibrium of FTO risk allele.
```{r check Hardy-Weinberg equilibrium, echo=FALSE}
#FTO
MAF=(table(final_sample[,"FTO_risk"])[2]+2*table(final_sample[,"FTO_risk"])[3])/(nrow(final_sample[,])*2)
HW.test <- HWChisq(table(final_sample[,"FTO_risk"]),verbose=TRUE)

#For Taq1A, we only provide the binarized variable in the synthetic dataset.
```

# Association of BMI and reward network connectivity
Here, we test for the linear association of BMI and reward network connectivity measures. If BMI is significant based on `anova` model comparison, we would perform the second analysis adjusting for covariates (not included in the synthetic dataset).

## FA CC ~ BMI 
```{r FA CC model, echo=FALSE}
##FA_CC
R1 <- lm(rel_FA_CC ~ Age_all + factor(sex) + BMI_BMI , data=final_sample)
R0 <- lm(rel_FA_CC ~ Age_all + factor(sex), data=final_sample)
anova(R1,R0) #p=0.02419 -> no adjusted model
summary(R1)
#lm.beta(R1)
confint(R1)
cbind(coefficients(R1),confint(R1))

##graphically show confidence intervalls
round(confint(R1),4)
##first check assumptions
#diagnostics-> normally distributed residuals with homogenous variance
hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))

plot(x=fitted(R1),y=residuals(R1))
plot(residuals(R1))
abline(a=0,b=0)

qqnorm(residuals(R1))
qqline(residuals(R1))

##model stability-> looks ok
max(abs(dffits(R1)))
head(dfbeta(R1))
xx=cbind(coefficients(R1),coefficients(R1)+t(apply(dfbeta(R1),MARGIN=2, FUN=range)))
colnames(xx)=c("orig","min","max")
round(xx,5)
max(as.vector(influence(R1)$hat))

##MULTIcollinearity
vif(R1) # largest vif > 10 is problematic
1/vif(R1) # Tolerance < 0.1 serious problem; <0.2 potential problem
mean(vif(R1)) # Average SUBSTANTIALLY > 1, the regression might be biased
###
```

## FA CS ~ BMI (Model 1)
```{r CS FA, echo=FALSE}
R1 <- lm(rel_FA_strength ~ Age_all + factor(sex) + BMI_BMI , data=final_sample)
R0 <- lm(rel_FA_strength ~ Age_all + factor(sex), data=final_sample)
anova(R1,R0)
summary(R1)
confint(R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)

```
## NOS CC ~ BMI (Model 1)
```{r NOS CC, echo=FALSE}
R1 <- lm(rel_NOS_CC ~ Age_all + factor(sex)+ BMI_BMI , data=final_sample)
R0 <- lm(rel_NOS_CC ~ Age_all + factor(sex), data=final_sample)
anova(R1,R0)
summary(R1)
confint(R1)
```
## NOS CS ~ BMI (Model 1)
```{r NOS strength, echo=FALSE}
R1 <- lm(rel_NOS_strength ~ Age_all + factor(sex) + BMI_BMI, data=final_sample)
R0 <- lm(rel_NOS_strength ~ Age_all + factor(sex) , data=final_sample)
anova(R0,R1)
summary(R1)
confint(R1)

hist(residuals(R1),probability = T)
x=seq(from=min(residuals(R1)),
      to=max(residuals(R1)), length.out=100)
lines(x=x,y=dnorm(x, mean=0, sd=sd(residuals(R1))))
qqPlot(R1)
```

# Association of genotypes and BMI
First, we tested for an interaction of FTO and Taq1A on BMI.
```{r Interaction of FTO and Taq1A, echo=FALSE}
R1 <- lm(BMI_BMI~ Age_all + factor(sex) + as.factor(FTO_risk)*as.factor(Taq1A_risk)  , data=final_sample)
R0 <- lm(BMI_BMI~ Age_all + factor(sex)+ as.factor(FTO_risk) + as.factor(Taq1A_risk), data=final_sample)
summary(R1)
anova(R1,R0)
```
Then we estimated the main effect of both genotypes.
```{r main effect of genotypes, echo=FALSE}
summary(R0)
```
Like in the original data, more FTO alleles predicted higher BMI. In the synthetic data, the association of carrying a Taq1A risk allele and lower BMI becomes significant. The same trend was visible in the real data.

# Association of genotypes and reward network connectivity
Finally, we checked for a link between obesity-related genotypes and reward network connectivity. We saw no significant interaction, like in the real data, between FTO and Taq1A genotypes. In the synthetic data, there is a trend towards higher FA CC and FA strength for Taq1A risk carriers, which we did not find in the real data. There was no significant effect of FTO genotype.

## genotype ~ FA CC
```{r genotype and reward network connectivity FA CC, echo=FALSE}
R1 <- lm(rel_FA_CC ~ Age_all + factor(sex) + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
  R0 <- lm(rel_FA_CC ~ Age_all + factor(sex) + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R1)
summary(R1)

summary(R0)
```
## genotype ~ FA CS
```{r genotype and reward network connectivity FA CS, echo=FALSE}
R1 <- lm(rel_FA_strength ~ Age_all + factor(sex) + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
  R0 <- lm(rel_FA_strength ~ Age_all + factor(sex) + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R1)
summary(R1)

summary(R0)
```
## genotype ~ NOS CC
```{r genotype and reward network connectivity NOS CC, echo=FALSE}
R1 <- lm(rel_NOS_CC ~ Age_all + factor(sex) + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
  R0 <- lm(rel_NOS_CC ~ Age_all + factor(sex) + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R1)
summary(R1)

summary(R0)
```
## genotype ~ NOS CS
```{r genotype and reward network connectivity NOS CS, echo=FALSE}
R1 <- lm(rel_NOS_strength ~ Age_all + factor(sex) + as.factor(Taq1A_risk)*as.factor(FTO_risk) , data=final_sample)
  R0 <- lm(rel_NOS_strength ~ Age_all + factor(sex) + as.factor(Taq1A_risk) + as.factor(FTO_risk) , data=final_sample)
anova(R0,R1)
summary(R1)

summary(R0)
```

## Exploratory analysis in normal-weight participants
As reported in the discussion, we found a trend toward higher BMI in normal-weight FTO carriers. We see a similar pattern in the synthetic data, though it is not significant in the model comparison.
```{r exploratory analysis in normal-weight participants, echo=FALSE}
final_sample_nw=final_sample[final_sample$BMI_BMI<=25&final_sample$BMI_BMI>=18,]
table(final_sample$BMI_BMI<=25&final_sample$BMI_BMI>=18)


R1 <- lm(rel_NOS_strength ~ Age_all + factor(sex) + as.factor(FTO_risk) , data=final_sample_nw)
R0<- lm(rel_NOS_strength ~ Age_all + factor(sex),data=final_sample_nw)
anova(R1,R0)
summary(R1)
```